---
title: "Overview of the functions of the FARS package"
author: "Aaron M. Gusdon"
date: "2017.04.23"
output: rmarkdown::fars_vignette
vignette: >
  %\VignetteIndexEntry{Overview of the FARS package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo=FALSE, include=FALSE}
library(fars)
library(dplyr)
library(maps)
```

This vignette gives an overview of teh FARS package and the purpose of each function.  Its functions analyze data from teh Fatality Analysis Reporting System (FARS) [[found here](http://www.nhtsa.gov/Data/Fatality-Analysis-Reporting-System-(FARS)]

## Package functions

There are three exported functions for users:

- 'fars_read()'
- 'fars_summarize_years()'
- 'fars_map_state()'

A brief description with examples will be provided for each function.  The FARS data for the examples can be found in the package directory under teh '\extdta' folder.

## Read FARS data files

The function `fars_read()` reads data from a file path provided as an argument. It returns a `tibble` with the loaded data:

```{r fars_read_example}
filename <- system.file("extdata/accident_2013.csv.bz2", package = "fars")
fars_read(filename)
```

## Summarize number of accidents

The function `fars_summarize_years()` takes a vector or list of years (numeric/integer values) as an argument and if the corresponding files with data are available, it returns a pivot table with number of accidents per year and month. The data files need to be located in the working directory.

```{r fars_summarize_years_example}
setwd(system.file("extdata", package = "fars"))
fars_summarize_years(2013:2015)
```

## Plot accident locations

The last function available in the package, `fars_map_state()` takes a state index and year as arguments and returns a plot of the state with accident locations. The state indices need to correspond available indeces from the data set. Here are some examples:

```{r fars_map_state_example}
setwd(system.file("extdata", package = "fars"))
fars_map_state(45, 2015)
```

The data files need to be located in the working directory